---
title: "Introduction to Meta-Analysis"
author: "Marta Mainetti"
date: "2026-05-21"
output: pdf_document
---

# Import libraries
```{r, message=FALSE, warning=FALSE}
library(dplyr)
library(meta)
library(readxl)
```

# Load data

```{r echo=FALSE}
AcuteManiaP_raw <- read_excel("/Users/mm25l304/Documents/GitHub/lecture-introduction-to-meta-analysis/Rcodes/AcuteManiaP.xls")
```


```{r}
AcuteManiaP <- AcuteManiaP_raw %>% filter(treat2 == "Risperidone")

str(AcuteManiaP)
```

```{r}
nrow(AcuteManiaP)
```

# Meta-analysis of a continuous outcome 

```{r}
pooledSMD1 <- metacont(mean.c = mean1, sd.c = sd1, n.c = ncont1,
                       mean.e = mean2, sd.e = sd2, n.e = ncont2,
                       data = AcuteManiaP, studlab = studlab, sm = "SMD")

summary(pooledSMD1)
```

5 studies were involved in the meta-analysis the fixed effect model gave an 
estimate of -0.60 [-0.73; -0.48] for the fixed effects and -0.59 [-0.76; -0.43] for the random effects. Both significant.

The test of heterogeneity does not highlight evidence of heterogeneity. This 
doesn't necessarily mean that there's no heterogeneity, it might be due to low power, i.e.
a small number of study. 

```{r}
summary(update(pooledSMD1, method.tau = "PM")) 
```

The random effects model now estimates that the treatment effect is -0.60 [-0.76; -0.43]

However, with only 5 studies, it's difficult to estimate the heterogeneity
(we need at least >5 studies). 

```{r}
summary(update(pooledSMD1, method.random.ci = T)) 
```

## Forest plots

```{r}
forest(pooledSMD1, sortvar = ncont1 + ncont2,
       prediction = T, common = T)
```

# Meta-analysis of a binary outcome

Odds Ratio as Effect Size

```{r}
pooledOR <- metabin(event.e = event2,
                    n.e = n2,
                    event.c = event1,
                    n.c = n1,
                    data = AcuteManiaP,
                    studlab = studlab, sm = "OR")
summary(pooledOR)
```

```{r}
forest(pooledOR, sortvar = n1 + n2, layout = "JAMA",  allstudies = F,
             fs.study = 10, ff.study = "italic",
             col.diamond = "indianred", col.diamond.lines = "dodgerblue",
             col.square = "salmon", col.square.lines = "salmon")
```

Risk Ratio as Effect Size

```{r}
summary(update(pooledOR, sm = "RR")) 
```














