---
title: "Introduction to Meta-Analysis"
author: "Marta Mainetti"
date: "2026-05-21"
output: pdf_document
---

# Import libraries
```{r, message=FALSE, warning=FALSE}
library(dplyr)
library(meta)
library(readxl)
```

# Load data

```{r echo=FALSE}
setwd("~/Desktop/unimi_lecture/script")

AcuteManiaP_raw <- read_excel("AcuteManiaP.xls")
```


```{r}
AcuteManiaP <- AcuteManiaP_raw %>% filter(treat2 == "Risperidone")

str(AcuteManiaP)
```

```{r}
nrow(AcuteManiaP)
```

```{r}
pooledSMD1 <- metacont(mean.c = mean1, sd.c = sd1, n.c = ncont1,
                       mean.e = mean2, sd.e = sd2, n.e = ncont2,
                       data = AcuteManiaP, studlab = studlab, sm = "SMD")

summary(pooledSMD1)
```

5 studies were involved in the meta-analysis the fixed effect model gave an 
estimate of -0.60 [-0.73; -0.48]  random effects model: -0.59 [-0.76; -0.43]

Both significant

The test of heterogeneity does not highlight evidence of heterogeneity. This 
doesn't mean that there's no heterogeneity, it might be low power

meta analysis: in this case I specify the estimator PM of tau (also the default)
and I ask not to compute a fixed-effect meta-analysis

```{r}
summary(update(pooledSMD1, method.tau = "PM")) 
```

random effects model: -0.60 [-0.76; -0.43]
It is very similar to REML case (default)
However, with only 5 studies, it's difficult to estimate the heterogeneity
(we need at least >5 studies). 

```{r}
summary(update(pooledSMD1, method.random.ci = T)) 
```

Hartung-Knapp adjustment for random effects model (df = 4)

# FOREST PLOTS

```{r}
forest(pooledSMD1, sortvar = ncont1 + ncont2,
       prediction = T, common = T)
```

# Meta-analysis with binary outcome
Odds-Ratio as Effect Size

```{r}
pooledOR <- metabin(event.e = event2,
                    n.e = n2,
                    event.c = event1,
                    n.c = n1,
                    data = AcuteManiaP,
                    studlab = studlab, sm = "OR")
summary(pooledOR)
```

```{r}
forest(pooledOR, sortvar = n1 + n2, layout = "JAMA",  allstudies = F,
             fs.study = 10, ff.study = "italic",
             col.diamond = "indianred", col.diamond.lines = "dodgerblue",
             col.square = "salmon", col.square.lines = "salmon")
```

```{r}
summary(update(pooledOR, sm = "RR")) 
```














